\LARGE \bf IMU-based pose-estimation for spherical robots with limited resources

### 2021-05-31 mfi21/ieeeconf/root.tex
minor improvements

To test the impact of the filter on 3D mapping, we used the prototype in Figure \ref{fig:prototype}.
After 2 s of initial standstill, the sphere rolled aprox. 4 m within 5 s in an indoor environment from one corner of the room to the opposing one.
We extracted the laserscan data, as well as the filtered pose using the presented algorithm.
Without further pre- or post-processing, the pose data is combined with the laserscan data to create a 3D map.
Figure \ref{fig:mapping} compares the result with another 3D map, which is generated the exact same way, but using a different pose filter.
Namely a combination of the Madgwick filter for orientation and a double integration of the accelerometer mean for position.
The result of this filter clearly got improved by the presented algorithm as it provides a better pose estimation.
The more detailed part of the map, as seen in the bottom left corner of the above image of Figure \ref{fig:mapping}, corresponds to the initial standstill.
It is more detailed than the rest of the map because of the non-repeating flower-shaped scanning pattern, where point density increases with time.
Note that the same detailed part is also present in the below image.
However, it is out of view (at the bottom right corner) as the Madgwick filter misses the initial pose estimation by a lot. 
The presented algorithm also performs better when the sphere is moving, as the walls and the ceiling are more recognizable, which is favourable for many SLAM algorithms.

>>>>

To test the impact of the filter on 3D mapping, we used the prototype in Figure \ref{fig:prototype}.
After 2 s of initial standstill, the sphere rolled aprox. 4 m within 5 s in an indoor environment from one corner of the room to the opposing one.
We extracted the laserscan data, as well as the filtered pose using the presented algorithm.
Without further pre- or post-processing, the pose data is combined with the laserscan data to create a 3D map.
Figure \ref{fig:mapping} compares the result with another 3D map, which is generated in the same fashion.
However, a combination of the Madgwick filter for orientation and a double integration of the accelerometer mean for position, is used.
The presented filter clearly improves the resulting map compared to the Madgwick/integrator approach.
In particular, the walls and the ceiling are more recognizable, which is favourable for many SLAM algorithms.
Another improvement is that points were no longer sensed below the ground, i.e. the plane on which the sphere was moving on.
The more detailed part of the map, as seen in the bottom left corner of the above image of Figure \ref{fig:mapping}, corresponds to the initial standstill.
It is more detailed than the rest of the map because of the non-repeating flower-shaped scanning pattern, where point density increases with time.
Note that the same detailed part is also present in the image below.
However, it is out of view (at the bottom right corner) as the Madgwick filter misses the initial pose estimation by a lot.
